The Development of a Collaborative Tool to Teach Debugging

### 2020-11-27 paper/paper.tex
Added some tool writeups

\documentclass[12pt]{article}
\usepackage{multicol}
\usepackage{cite}
\usepackage{listings}
\usepackage{tikz}
\usepackage[doublespacing]{setspace}

\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}
\usetikzlibrary{positioning}

\lstset{basicstyle=\ttfamily}

@@

\subsubsection{Mozilla's rr}

@@

rr is ``a lightweight tool for recording, replaying and debugging
execution of applications''.\cite{rr-repo} rr allows a programmer to
record the execution of a program on any compatible machine and replay
the execution later.  This enhances GDB's ability to ``time-travel''
when debugging, using commands such as \lstinline{reverse-continue}
and \lstinline{reverse-stepi}\cite{gdbman} to step backwards and
forwards through a program's execution.  Through a novel encapsulation
of the execution space, rr is able to deterministically record and
replay the execution of syscalls and other process behavior that
differs run-to-run.  This is invaluable when trying to debug behavior
that is not entirely dependent on the code being debugged.  A typical
workflow in rr consists of recording an inexplicable error, replaying
execution to find the area in which the error occurs, and then
narrowing in on the bug not by re-running the entire program, but by
progressing back and forth through execution in the problem area.

@@

In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
that rely on capturing the entire state of of a virtual machine to
replay execution, rr records and replays faster, produces far smaller
files, and doesn't force execution inside of a
VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The tradeoff for these
benefits are two major system limitations: rr is only compatible with
the Linux kernel, and it's deterministic recording and replay relies
on a feature that is only found on modern \textit{Intel} x86 CPUs.
These limitations influenced the development of this project as a
webapp similar to existing tools for collaborative programming.
Luckily, the speed and size benefits of rr lend themselves well to
non-local execution.

\subsubsection{Kubernetes and Docker}

TODO: Write about k8s and docker.

\subsubsection{React}

TODO: Write about react.

\subsubsection{Xterm.js}

TODO: Write about Xterm.js if I end up using it (which seems
overwhelmingly likely ATM).

>>>>

\documentclass[12pt]{article}
\usepackage{multicol}
\usepackage{cite}
\usepackage{listings}
\usepackage{graphicx}
\usepackage[doublespacing]{setspace}
\usepackage{hyperref}
\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}

\lstset{basicstyle=\ttfamily}
\graphicspath{{./images/}}

@@

\begin{figure}[h!]

  \includegraphics[scale=.8]{overall_system}
  \centering
  \caption{Overview of the Collaborative Debugger}
  \label{debugger:overview}
\end{figure}

The next sections give an overview of the various tools used to create
the collaborative debugger.  The debugger consists of:

\begin{enumerate}
\item A frontend web app built using React (\ref{react}) that presents a debugging
  interface to the end user.
\item A distributed backend managed by Kubernetes (\ref{k8s}), split into three parts:
  \begin{enumerate}
  \item A pod for each debugging instance which runs the rr debugger
    (\ref{rr}).  These communicate directly with users through
    WebSocket server pods. (\ref{socketio}).
  \item Pods running a frontend server written in Node which works in
    tandem with an API server created using Flask (\ref{flask/node}).
    The API server manages creation and destruction of debug sessions,
    as well as authentication.
  \item A MongoDB database.
  \end{enumerate}
\end{enumerate}

\subsubsection{Kubernetes}\label{k8s}

Kubernetes is the defacto standard in container orchestration
software.  It provides a layer of abstraction on top of normal
containers, like those created by Docker.  By bundling one or more
closely linked containers into a ``pod'', Kubernetes is able to manage
deployment and re-deployment of applications running inside
containers.  It is trivial to create new pods, or to create multiple
pods running the same application as needed within a Kubernetes
cluster \cite{k8s}.  The speed at which even relatively large pods can be created
and the inherent security provided by containerization drove the
decision to create a new pod on the fly for each debugging instance in
the collaborative debugger.
\par

Kubernetes also provides services to facilitate load balancing, manage
storage volumes, and contain secrets.  The abstraction provided by
these features, in tandem with the ease of Kubernetes deployment on a
managed Kubernetes service\cite{do_managed_k8s} greatly accelerated
development.

\subsubsection{Mozilla's rr}\label{rr}

@@

rr is ``a lightweight tool for recording, replaying and debugging
execution of applications''\cite{rr-repo}. rr allows a programmer to
record the execution of a program on any compatible machine and replay
the execution later.  This enhances GDB's ability to ``time-travel''
when debugging, using commands such as \lstinline{reverse-continue}
and \lstinline{reverse-stepi}\cite{gdbman} to step backwards and
forwards through a program's execution.  Through a novel encapsulation
of the execution space, rr is able to deterministically record and
replay the execution of syscalls and other process behavior that
differs run-to-run.  This is invaluable when trying to debug behavior
that is not entirely dependent on the code being debugged.  A typical
workflow in rr consists of recording an inexplicable error, replaying
execution to find the area in which the error occurs, and then
narrowing in on the bug not by re-running the entire program, but by
progressing back and forth through execution in the problem area.
\par

rr is an ideal tool for teaching debugging because it allows
instructors to record execution of a program and design a debugging
example with the knowledge that normally non-deterministic events will
be repeatable, and that any input they provide to the program will be
exactly replicated.  With the collaborative debugger, teachers can
record a program's execution and design a debugging lesson which
students can work on together.  The repeatablity of rr means that
students can focus on debugging, and teachers can create as specific
examples as they please.

@@

In comparison to solutions like PANDA\cite{10.1145/2843859.2843867}
that rely on capturing the entire state of of a virtual machine to
replay execution, rr records and replays faster, produces far smaller
files, and doesn't force execution inside of a
VM.\cite{DBLP:journals/corr/OCallahanJFHNP17} The tradeoff for these
benefits are two major system limitations: rr is only compatible with
the Linux kernel, and it's deterministic recording and replay relies
on a feature that is only found on modern \textit{Intel} x86 CPUs.
These limitations influenced the development of this project as a
webapp similar to existing tools for collaborative programming.
\par

Luckily, the speed and size benefits of rr lend themselves well to
non-local execution.  In conjunction with Kubernetes, it takes a few
seconds to create a new container running rr and connect to web
clients.

\subsubsection{pygdbmi}

\subsubsection{Socket.IO}\label{socketio}

To speed communication, the collaborative debugger uses WebSockets to
directly connect web clients and the pods running rr.  Socket.IO is a
library that extends WebSockets.  It provides backup incase a
WebSocket connection cannot be established, enables automatic
reconnection and disconnection detection, and adds support for
namespaces \cite{socketio}.  The collaborative debugger uses the
standard Javascript implementation of Socket.IO on the client side.
Messages are passed through a server to individual debugging pods,
both of which use the Python implementation of Socket.IO,
python-socket.io \cite{python_socketio}.

\subsubsection{Flask and Node}\label{flask/node}

\subsubsection{React and Xterm.js}\label{react}
